---
title: "ETF5231 Group Assignment 3"
subtitle: 'Group7 GenieUS'
author:
  - Ting-Ting Wu (35717343)
  - Tzu-Hsuan, Yang (35461586)
  - Nutchanok Saitin (35493011)
  - Heng-Hsieh, Chang (35567201)
format:
  pdf
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(echo = TRUE, eval=TRUE, error=TRUE, cache=FALSE, warning=FALSE, message=FALSE, fig.width = 10)

library(fpp3)
library(readr)
library(dplyr)
library(ggplot2)
library(patchwork)
library(knitr)
library(kableExtra)
```

```{r}
#| include: false

meta <- read_csv("GA_data.csv", col_names = TRUE, n_max = 3)

dat <- read_csv("GA_data.csv", 
                col_names = colnames(meta),
                skip = 4)

print_retail_act <- dat %>% 
  rename(Month = "Student ID", y ="Group 7") %>%
  select(Month, y) %>% 
  mutate(Month=yearmonth(Month)) %>% 
  as_tsibble(index = Month)
```

For the tasks that follow you will be modelling the original data without any transformation performed (even if you previously deemed it necessary). 


## Question 1

*Plot your time series. By observing the plot and describing its components select an ETS model you think is appropriate for forecasting. Make sure you justify your choice (no more than 150 words). (8 marks)*

As shown in @fig-q1-timeplot, the ACT newspaper and book retailing series has clear monthly fluctuations and a changing overall level. The series rises in the earlier years, stays relatively high around the middle of the sample, and then falls strongly in the later years. Because the direction of the series changes over time, we do not think an additive or damped trend is suitable. The size of the seasonal fluctuations and the random variation also seem larger when turnover is high, and smaller when turnover is low. This means the errors are more likely to be proportional to the level, so a multiplicative error structure is reasonable. Therefore, we choose ETS(M,N,M), which allows multiplicative errors and seasonality without forcing a fixed trend.

```{r}
#| label: fig-q1-timeplot
#| fig-cap: "Time plot of monthly newspaper and book retail turnover in ACT (1982-2023)."

print_retail_act %>% 
  autoplot(y) +
  labs(x = "Month",
       y = "Turnover (Millions of Dollars)", 
       title = "Newspaper and book retailing",
       subtitle = "Australian Capital Territory")
```

\newpage
## Question 2

*Search for an ETS model for your series using appropriate information criteria. Describe the process you have implemented. Which part of the data is used in this process? Which model have you selected and why? Show all necessary output (10 marks)*

The `ETS()` function considers the admissible ETS models and selects the model with the lowest AICc, balancing goodness of fit and model complexity. The automatic search selected ETS(M,N,M), which is consistent with our choice in Question 1. The whole available time series was used in this process, because this question focuses on information criteria rather than out-of-sample evaluation. The output from `report()` shows that the selected model has multiplicative errors, no trend and multiplicative seasonality. The reported information criteria are $AIC = 2744.476$, $AICc = 2745.478$ and $BIC = 2807.544$. This supports our visual choice, because both the time plot and the AICc-based search suggest proportional errors and seasonal variation, but no stable long-term trend.

```{r}
#| label: tbl-q2-auto-report
#| tbl-cap: "Automatically selected ETS model for the Group 7 series."

auto_fit <- print_retail_act %>%
  model(auto_ets = ETS(y))

report(auto_fit)
```

```{r}
#| label: fig-q2-components
#| fig-cap: "Estimated components from the ETS(M,A,M) model."
#| fig-width: 10
#| fig-height: 10

auto_fit %>%
  components() %>%
  autoplot()
```

\newpage
## Question 3

*Now use time series cross-validation to compare the model selected from the previous step, along with ETS(M,N,M), ETS(M,A,M) and ETS(M,Ad,M) models. You will need to think about more than h=1 step-ahead. Think about an appropriate range of forecast horizons and justify your choice. Fully describe the process you have implemented. Show all necessary output. (Hint the code at the end of this section [https://otexts.com/fpp3/tscv.html](https://otexts.com/fpp3/tscv.html) may help you to get started.) (22 marks)*

Time series cross-validation was used to compare the candidate ETS models beyond one-step-ahead forecasts. We used 12-month and 24-month forecast horizons, representing one and two full seasonal cycles. For both horizons, `.init = 180` was used, giving 15 years of initial monthly data, enough to estimate the monthly seasonal pattern. In both cases, `.step = 1` was used so every possible monthly forecast origin after the initial training period was used.

The 12-month results in @tbl-q3-cv-accuracy-12 show that ETS(M,N,M) has the lowest RMSE, MAE, MAPE, MASE and RMSSE, although ETS(M,Ad,M) is very close. In @fig-q3-rmse-h12, RMSE increases as the forecast horizon increases. The three models are close at early horizons, but ETS(M,A,M) becomes worse at longer horizons, while ETS(M,N,M) and ETS(M,Ad,M) remain similar.

The 24-month results in @tbl-q3-cv-accuracy-24 show the same ranking, with ETS(M,N,M) still slightly better overall. In @fig-q3-rmse-h24, RMSE again increases with horizon. ETS(M,A,M) moves further away from the other two models, while ETS(M,N,M) and ETS(M,Ad,M) almost overlap for much of the horizon. This suggests that a damped trend gives similar longer-horizon forecasts, but does not clearly improve accuracy.

Overall, the cross-validation results support selecting ETS(M,N,M). It performs slightly best for both horizons, so the simpler no-trend model is preferred over the additive or damped trend alternatives.

```{r}
#| label: tbl-q3-cv-accuracy-12
#| tbl-cap: "Time series cross-validation accuracy for the candidate ETS models using a 12-month forecast horizon."

print_retail_cv <- print_retail_act |>
  stretch_tsibble(.init = 180, .step = 1)

cv_fit <- print_retail_cv |>
  model(mnm = ETS(y ~ error("M") + trend("N")  + season("M")),
        mam = ETS(y ~ error("M") + trend("A")  + season("M")),
        madm = ETS(y ~ error("M") + trend("Ad") + season("M")))

cv_fc_12 <- cv_fit |>
  forecast(h = 12)

cv_acc_12 <- cv_fc_12 |>
  accuracy(print_retail_act) |>
  arrange(RMSE)

cv_acc_12 |>
  select(.model, .type, RMSE, MAE, MAPE, MASE, RMSSE, ACF1) |>
  kable(digits = 3)
```

```{r}
#| label: fig-q3-rmse-h12
#| fig-cap: "Cross-validation RMSE by forecast horizon for the candidate ETS models using a 12-month forecast horizon."

cv_fc_12_h <- cv_fc_12 |>
  group_by(.id, .model) |>
  mutate(h = row_number()) |>
  ungroup() |>
  as_fable(response = "y", distribution = y)

cv_fc_12_h |>
  accuracy(print_retail_act, by = c("h", ".model")) |>
  ggplot(aes(x = h, y = RMSE, colour = .model)) +
  geom_point() +
  geom_line() +
  scale_x_continuous(breaks = seq(1, 12, by = 2)) +
  labs(x = "Forecast horizon", y = "RMSE",
       colour = "Model",
       title = "Cross-validation RMSE by forecast horizon",
       subtitle = "12-month forecast horizon") +
  theme_minimal()
```

```{r}
#| label: tbl-q3-cv-accuracy-24
#| tbl-cap: "Time series cross-validation accuracy for the candidate ETS models using a 24-month forecast horizon."

cv_fc_24 <- cv_fit |>
  forecast(h = 24)

cv_acc_24 <- cv_fc_24 |>
  accuracy(print_retail_act) |>
  arrange(RMSE)

cv_acc_24 |>
  select(.model, .type, RMSE, MAE, MAPE, MASE, RMSSE, ACF1) |>
  kable(digits = 3)
```

```{r}
#| label: fig-q3-rmse-h24
#| fig-cap: "Cross-validation RMSE by forecast horizon for the candidate ETS models using a 24-month forecast horizon."

cv_fc_24_h <- cv_fc_24 |>
  group_by(.id, .model) |>
  mutate(h = row_number()) |>
  ungroup() |>
  as_fable(response = "y", distribution = y)

cv_fc_24_h |>
  accuracy(print_retail_act, by = c("h", ".model")) |>
  ggplot(aes(x = h, y = RMSE, colour = .model)) +
  geom_point() +
  geom_line() +
  scale_x_continuous(breaks = seq(1, 24, by = 2)) +
  labs(x = "Forecast horizon", y = "RMSE", 
       colour = "Model", 
       title = "Cross-validation RMSE by forecast horizon", 
       subtitle = "24-month forecast horizon") +
  theme_minimal()
```

\newpage
## Question 4

*Which model do you prefer for your data? Choose a model and briefly explain your choice. Plot and describe its components and comment on the estimated parameters and how these are appropriate for your data. Perform all the necessary diagnostic checks and comment. (max 400 words) (14 marks)*

```{r}

```

\newpage
## Question 5

*Generate forecasts for the two years following the end of your sample using your chosen/best model. Plot them and briefly comment on these. (No more than 100 words). (4 marks)*

```{r}

```