---
title: "ETF5231 Group Assignment 3"
subtitle: 'Group7 GenieUS'
author:
  - Ting-Ting Wu (35717343)
  - Tzu-Hsuan, Yang (35461586)
  - Nutchanok Saitin (35493011)
  - Heng-Hsieh, Chang (35567201)
format:
  pdf
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(echo = TRUE, eval=TRUE, error=TRUE, cache=FALSE, warning=FALSE, message=FALSE, fig.width = 10)

library(fpp3)
library(readr)
library(dplyr)
library(ggplot2)
library(patchwork)
library(knitr)
library(kableExtra)
```

```{r}
#| include: false

meta <- read_csv("GA_data.csv", col_names = TRUE, n_max = 3)

dat <- read_csv("GA_data.csv", 
                col_names = colnames(meta),
                skip = 4)

print_retail_act <- dat %>% 
  rename(Month = "Student ID", y ="Group 7") %>%
  select(Month, y) %>% 
  mutate(Month=yearmonth(Month)) %>% 
  as_tsibble(index = Month)
```

For the tasks that follow you will be modelling the original data without any transformation performed (even if you previously deemed it necessary). 


## Question 1

*Plot your time series. By observing the plot and describing its components select an ETS model you think is appropriate for forecasting. Make sure you justify your choice (no more than 150 words). (8 marks)*

```{r}

```

\newpage
## Question 2

*Search for an ETS model for your series using appropriate information criteria. Describe the process you have implemented. Which part of the data is used in this process? Which model have you selected and why? Show all necessary output (10 marks)*

```{r}

```
\newpage
## Question 3

*Now use time series cross-validation to compare the model selected from the previous step, along with ETS(M,N,M), ETS(M,A,M) and ETS(M,Ad,M) models. You will need to think about more than h=1 step-ahead. Think about an appropriate range of forecast horizons and justify your choice. Fully describe the process you have implemented. Show all necessary output. (Hint the code at the end of this section [https://otexts.com/fpp3/tscv.html](https://otexts.com/fpp3/tscv.html) may help you to get started.) (22 marks)*

```{r}

```

\newpage
## Question 4

*Which model do you prefer for your data? Choose a model and briefly explain your choice. Plot and describe its components and comment on the estimated parameters and how these are appropriate for your data. Perform all the necessary diagnostic checks and comment. (max 400 words) (14 marks)*

```{r}

```

\newpage
## Question 5

*Generate forecasts for the two years following the end of your sample using your chosen/best model. Plot them and briefly comment on these. (No more than 100 words). (4 marks)*

```{r}

```