---
title: "ETF5231 Group Assignment 4"
subtitle: 'Group7 GenieUS'
author:
  - Ting-Ting Wu (35717343)
  - Tzu-Hsuan, Yang (35461586)
  - Nutchanok Saitin (35493011)
  - Heng-Hsieh, Chang (35567201)
format: pdf
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=TRUE, error=TRUE, cache=TRUE, warning = FALSE)
# write the appropriate code to read in your group data from Group_Data.csv

library(fpp3)
library(readr)
library(dplyr)
library(ggplot2)
library(patchwork)
library(knitr)
library(kableExtra)
library(readxl)
```

## Question 1

*Update your data from the ABS website and plot it. Do you see a COVID-19 effect? (3 marks)*。
```{r}
#| label: update-dataprep

extended_dat <- read_excel("GA4_data.xlsx", sheet = "Data1", skip = 9) 

print_retail_act <- extended_dat |>
  rename(Month = "Series ID", y = "A3349773T") |>
  select(Month, y) |>
  mutate(Month = yearmonth(Month)) |>
  as_tsibble(index = Month)
```

```{r}
#| label: fig-extended-series
#| fig-width: 10
#| fig-height: 8
#| fig-cap: "Updated turnover series from Apr 1982 to Jun 2025. The shaded area highlights March 2020 to Sep 2022, when ABS did not publish trend estimates due to COVID-related disruption and volatility. The left plot shows the whole series, while the right plot zooms in from Jan 2018 to show recent movements."

plot_whole_series <- print_retail_act |>
  autoplot(y) +
  annotate(
    "rect",
    xmin = yearmonth("2020 Mar"),
    xmax = yearmonth("2022 Sep"),
    ymin = -Inf,
    ymax = Inf,
    alpha = 0.2,
    fill = "grey30"
  ) +
  labs(
    x = "Month",
    y = "Turnover (Millions of Dollars)"
  ) +
    theme_minimal()

plot_zoomed <- print_retail_act |>
  filter(Month >= yearmonth("2018 Jan")) |>
  autoplot(y) +
  annotate(
    "rect",
    xmin = yearmonth("2020 Mar"),
    xmax = yearmonth("2022 Sep"),
    ymin = -Inf,
    ymax = Inf,
    alpha = 0.2,
    fill = "grey30"
  ) +
  labs(
    x = "Month",
    y = "Turnover (Millions of Dollars)"
  ) +
    theme_minimal()

(plot_whole_series / plot_zoomed) +
  plot_annotation(
    title = "Turnover Series with COVID-19 Period Highlighted",
    subtitle = "Newspaper and book retailing, Australian Capital Territory"
  )
```

@fig-extended-series shows a long-term decline in turnover before COVID-19. During the COVID-disruption period, turnover remained low and volatile, with some sharp temporary movements, including the lowest point in the series. After the COVID-disruption period, turnover shows a slight recovery, but remains at a relatively low level, with the seasonal pattern still visible.

```{r}
#| label: fig-season-plot
#| fig-height: 4

print_retail_act |>
  filter(Month >= yearmonth("2018 Jan")) |>
  gg_season(y, labels = "both") +
  labs(
    title = "Turnover (Jan 2018 - Jun 2025)",
    subtitle = "Newspaper and book retailing, Australian Capital Territory",
    x = "Month",
    y = "Turnover (Millions of Dollars)"
  ) +
    theme_minimal()
```

@fig-season-plot shows that turnover generally declined over time while retaining a consistent seasonal pattern, with higher values toward the end of the year, especially in December. During the COVID-disruption period, the 2021 line falls sharply from July to September and reaches the lowest point, suggesting a strong temporary weakening. After the COVID-disruption period, turnover increases slightly in some months, but remains relatively low compared with 2018–2020.

Overall, COVID-19 appears to have contributed to additional volatility and a temporary weakening around 2021, although the series was already following a long-term downward trend before the disruption period.

\newpage
## Question 2

*Identify the best possible ARIMA model for your data and generate forecasts. Present a complete analysis and justifications for each step you take. There is no word limit here. No more hints will be provided - we leave it up to you to provide a full analysis. Treat this as a consulting job for a client wanting some ARIMA forecasts (20 marks)*

### Transformation

```{r}
#| label: tbl-q2-lambda
#| tbl-cap: "Guerrero lambda for the updated series."

lambda_tbl <- print_retail_act |>
  features(y, guerrero)

lambda <- lambda_tbl$lambda_guerrero

lambda_tbl |>
  mutate(lambda_guerrero) |>
  rename(`Guerrero lambda` = lambda_guerrero) |>
  kable(digits = 3)
```

```{r}
#| label: fig-q2-transformations
#| fig-cap: "Comparison of transformations for the updated turnover series."
#| fig-width: 10
#| fig-height: 16

p1 <- print_retail_act |>
  autoplot(y) +
  labs(title = "Original", x = "Month", y = "Turnover")

p2 <- print_retail_act |>
  autoplot(sqrt(y)) +
  labs(title = "Square-root", x = "Month", y = "sqrt(Turnover)")

p3 <- print_retail_act |>
  autoplot(log(y)) +
  labs(title = "Log", x = "Month", y = "log(Turnover)")

p4 <- print_retail_act |>
  autoplot(box_cox(y, lambda)) +
  labs(title = "Box-Cox using Guerrero lambda",
       x = "Month", y = "Box-Cox(Turnover)")

(p1 / p2 / p3 / p4)
```

@fig-q2-transformations compares the original, square-root, log and Box-Cox transformations. The original series has changing variation over time, so a transformation is needed to stabilise the variance before ARIMA modelling. Guerrero's lambda is around 0.424, meaning log is too strong and the original scale is not stable enough. Square-root is a reasonable approximation, but Box-Cox uses the estimated lambda directly, so I use `box_cox(y, lambda)`.

```{r}
#| label: fig-q2-boxcox
#| fig-cap: "Box-Cox transformed turnover series."
#| fig-width: 10
#| fig-height: 4

print_retail_act |>
  autoplot(box_cox(y, lambda)) +
  labs(title = "Box-Cox transformed turnover series",
       x = "Month",
       y = "Box-Cox(Turnover)")
```

### Differencing and stationarity

@fig-q2-boxcox still shows a long-run downward movement and regular monthly seasonality. Therefore, differencing is needed before selecting ARIMA orders.

```{r}
#| label: fig-q2-seasonal-difference
#| fig-cap: "Seasonally differenced Box-Cox transformed turnover series."
#| fig-width: 10
#| fig-height: 4

print_retail_act |>
  autoplot(difference(box_cox(y, lambda), 12)) +
  labs(title = "Seasonally differenced Box-Cox transformed turnover",
       x = "Month",
       y = "Seasonal difference of Box-Cox(Turnover)")
```

@fig-q2-seasonal-difference shows that seasonal differencing removes much of the repeated monthly pattern, but the series still has long swings above and below zero. This suggests that the transformed series may not yet be fully stationary, so I check whether an additional first difference is needed.

```{r}
#| label: fig-q2-seasonal-first-difference
#| fig-cap: "Seasonally and first differenced Box-Cox transformed turnover series."
#| fig-width: 10
#| fig-height: 4

print_retail_act |>
  autoplot(difference(box_cox(y, lambda), 12) |> difference()) +
  labs(title = "Seasonally and first differenced Box-Cox transformed turnover",
       x = "Month",
       y = "Differenced Box-Cox(Turnover)")
```

@fig-q2-seasonal-first-difference shows that after applying both seasonal and ordinary differencing, the series fluctuates more closely around zero with no clear remaining trend or seasonal pattern. Although some temporary shocks remain, this looks more stationary than using seasonal differencing alone. I therefore use one seasonal difference and one first difference, so $D = 1$, $d = 1$, and $m = 12$.

```{r}
#| label: tbl-q2-difference-tests
#| tbl-cap: "Unit root checks used as supporting evidence for differencing."

seasonal_diff_tbl <- print_retail_act |>
  features(box_cox(y, lambda), unitroot_nsdiffs)

first_diff_tbl <- print_retail_act |>
  features(difference(box_cox(y, lambda), 12), unitroot_ndiffs)

bind_cols(seasonal_diff_tbl, first_diff_tbl) |>
  kable()
```

@tbl-q2-difference-tests also supports this choice, suggesting one seasonal difference and one ordinary difference.

```{r}
#| label: tbl-q2-kpss
#| tbl-cap: "KPSS tests for the transformed series after differencing."

kpss_seasonal <- print_retail_act |>
  features(difference(box_cox(y, lambda), 12), unitroot_kpss) |>
  mutate(series = "After seasonal difference only")

kpss_final <- print_retail_act |>
  features(difference(box_cox(y, lambda), 12) |> difference(), unitroot_kpss) |>
  mutate(series = "After seasonal and first difference")

bind_rows(kpss_seasonal, kpss_final) |>
  select(series, kpss_stat, kpss_pvalue) |>
  mutate(across(where(is.numeric), ~ round(.x, 3))) |>
  kable()
```

@tbl-q2-kpss provides an additional stationarity check. Since the KPSS null hypothesis is stationarity, the seasonally differenced series is still not stationary at the 5% level ($p = 0.019$). After both seasonal and first differencing, I fail to reject stationarity ($p = 0.100$), supporting $D = 1$, $d = 1$, and $m = 12$.

### ARIMA model selection and diagnostics

```{r}
#| label: fig-q2-acf-pacf
#| fig-cap: "ACF and PACF of the stationary transformed series."
#| fig-width: 10
#| fig-height: 8

print_retail_act |>
  gg_tsdisplay(difference(box_cox(y, lambda), 12) |> difference(),
               plot_type = "partial",
               lag_max = 60) +
  labs(title = "Stationary transformed series after seasonal and first differencing")
```

@fig-q2-acf-pacf shows the ACF and PACF after the Box-Cox transformation, seasonal differencing and first differencing. The ACF has a strong spike at lag 1 and at the seasonal lag 12, suggesting non-seasonal MA(1) and seasonal MA(1). The PACF also shows early spikes up to about lag 2, so AR(2) is a plausible alternative. I therefore use ARIMA(0,1,1)(0,1,1)$_{12}$ as the initial model and compare it with plausible alternatives using AICc.

```{r}
#| label: q2-fit-arima-candidates

q2_arima_candidates <- print_retail_act |>
  model(
    arima011011 = ARIMA(box_cox(y, lambda) ~ pdq(0,1,1) + PDQ(0,1,1)),
    arima210011 = ARIMA(box_cox(y, lambda) ~ pdq(2,1,0) + PDQ(0,1,1)),
    arima011110 = ARIMA(box_cox(y, lambda) ~ pdq(0,1,1) + PDQ(1,1,0)),
    arima011210 = ARIMA(box_cox(y, lambda) ~ pdq(0,1,1) + PDQ(2,1,0)),
    arima011310 = ARIMA(box_cox(y, lambda) ~ pdq(0,1,1) + PDQ(3,1,0)),
    arima210110 = ARIMA(box_cox(y, lambda) ~ pdq(2,1,0) + PDQ(1,1,0)),
    arima210210 = ARIMA(box_cox(y, lambda) ~ pdq(2,1,0) + PDQ(2,1,0)),
    arima210310 = ARIMA(box_cox(y, lambda) ~ pdq(2,1,0) + PDQ(3,1,0))
  )
```

```{r}
#| label: tbl-q2-arima-ic
#| tbl-cap: "Information criteria for candidate ARIMA models."

q2_arima_candidates |>
  glance() |>
  arrange(AICc) |>
  select(.model, sigma2, log_lik, AIC, AICc, BIC) |>
  mutate(across(where(is.numeric), ~ round(.x, 3))) |>
  kable()
```

Based on @tbl-q2-arima-ic, ARIMA(0,1,1)(0,1,1)$_{12}$ has the lowest AICc, so it is the best candidate model among those considered. The next closest model is ARIMA(2,1,0)(0,1,1)$_{12}$, but its AICc is slightly higher. Both best-performing models include seasonal MA(1), while the models using seasonal AR terms have much higher AICc values. This suggests that the seasonal MA(1) structure is more appropriate than the seasonal AR alternatives. 

I therefore treat ARIMA(0,1,1)(0,1,1)$_{12}$ as the best manual candidate model and check its residual diagnostics before making the final model decision.

```{r}
#| label: q2-manual-arima

q2_manual_arima <- q2_arima_candidates |>
  select(arima011011)

q2_manual_arima |>
  report()
```

The estimated MA coefficient is -0.3249 and the seasonal MA coefficient is -0.6100. Both are negative, suggesting that recent non-seasonal and seasonal shocks are partly corrected in the following periods. This model is simple and is also consistent with the ACF pattern in @fig-q2-acf-pacf.

```{r}
#| label: fig-q2-manual-residuals
#| fig-cap: "Residual diagnostics for the manual ARIMA(0,1,1)(0,1,1)[12] candidate model."
#| fig-width: 10
#| fig-height: 8

q2_manual_arima |>
  gg_tsresiduals(lag = 48)
```

@fig-q2-manual-residuals shows that the residual time plot has no obvious remaining trend or seasonal pattern. The histogram is roughly symmetric and reasonably bell-shaped, although not perfectly normal. The residual ACF still has some significant spikes, so the residuals are not completely clean visually. Therefore, I use the Ljung–Box test as a formal check for remaining autocorrelation.

```{r}
#| label: tbl-q2-manual-ljung
#| tbl-cap: "Ljung-Box test for residual whiteness of the manual ARIMA candidate model."

q2_manual_arima |>
  augment() |>
  features(.innov, ljung_box, lag = 24, dof = 2) |>
  kable(digits = 3)
```

For @tbl-q2-manual-ljung, I use lag 24 and dof 2, from one non-seasonal MA term and one seasonal MA term. The null hypothesis is $H_0: \rho_1 = \cdots = \rho_{24} = 0$ jointly, meaning the residual autocorrelations up to lag 24 are jointly insignificant. Since $p = 0.001 < 0.05$, I reject $H_0$ at the 5% level. This suggests that the residuals are not white noise and that some autocorrelation remains. Therefore, I review the model rather than using this candidate model directly as the final forecasting model.

```{r}
#| label: q2-auto-arima

q2_auto_arima <- print_retail_act |>
  model(auto_arima = ARIMA(box_cox(y, lambda),
                           stepwise = FALSE,
                           approximation = FALSE))

q2_auto_arima |>
  report()
```

The manual candidate model failed the Ljung–Box test, so I use a full automatic ARIMA search with `stepwise = FALSE` and `approximation = FALSE`. The automatic search selects ARIMA(3,1,1)(1,1,1)$_{12}$ with the same Box-Cox transformation. It also uses $d = 1$ and $D = 1$, which agrees with the earlier differencing decision.

```{r}
#| label: tbl-q2-manual-auto-comparison
#| tbl-cap: "Information criteria comparison between the manual candidate ARIMA model and the automatic ARIMA model."

q2_compare_arima <- print_retail_act |>
  model(manual_arima = ARIMA(box_cox(y, lambda) ~ pdq(0,1,1) + PDQ(0,1,1)),
        auto_arima = ARIMA(box_cox(y, lambda),
                           stepwise = FALSE,
                           approximation = FALSE))

q2_compare_arima |>
  glance() |>
  arrange(AICc) |>
  select(.model, sigma2, log_lik, AIC, AICc, BIC) |>
  mutate(across(where(is.numeric), ~ round(.x, 3))) |>
  kable()
```

@tbl-q2-manual-auto-comparison shows that the automatic ARIMA model has a lower AICc than the manual candidate model. Since both models use the same Box-Cox transformation and differencing orders, the AICc comparison is appropriate. The automatic model therefore appears better, but I still check its residual diagnostics before using it for forecasting.

```{r}
#| label: fig-q2-auto-residuals
#| fig-cap: "Residual diagnostics for the automatic ARIMA(3,1,1)(1,1,1)[12] model."
#| fig-width: 10
#| fig-height: 8

q2_auto_arima |>
  gg_tsresiduals(lag = 48)
```

@fig-q2-auto-residuals shows the residual diagnostics for the automatic ARIMA(3,1,1)(1,1,1)$_{12}$ model. The residual time plot has no obvious trend or seasonal pattern, and the histogram is roughly symmetric and reasonably bell-shaped. However, the residual ACF still has a few significant spikes, so I use the Ljung–Box test as a formal check for remaining autocorrelation.

```{r}
#| label: tbl-q2-auto-ljung
#| tbl-cap: "Ljung-Box test for residual whiteness of the automatic ARIMA model."

q2_auto_arima |>
  augment() |>
  features(.innov, ljung_box, lag = 24, dof = 6) |>
  kable(digits = 3)
```

For @tbl-q2-auto-ljung, I use lag 24 and dof 6, from three non-seasonal AR terms, one non-seasonal MA term, one seasonal AR term and one seasonal MA term. The null hypothesis is $H_0: \rho_1 = \cdots = \rho_{24} = 0$ jointly. Since $p = 0.001 < 0.05$, I reject $H_0$ at the 5% level. This suggests that some residual autocorrelation remains, so the automatic model is not perfect.

**Final Decision**

The manual model is simpler, but I select the automatic ARIMA(3,1,1)(1,1,1)$_{12}$ model because it has the lower AICc. This is important because AICc already penalises the extra parameters, so the lower AICc suggests that the additional terms improve the fit. However, the Ljung–Box test still rejects white noise, so the model is not perfect and the forecasts should be interpreted cautiously.

### Forecasts from the final ARIMA model

```{r}
#| label: q2-final-forecast

q2_final_arima <- q2_auto_arima

q2_arima_fc <- q2_final_arima |>
  forecast(h = "24 months")

forecast_start <- max(print_retail_act$Month) + 1
```

```{r}
#| label: fig-q2-arima-forecast-full
#| fig-cap: "Twenty-four month forecasts from the final ARIMA model over the full updated series."
#| fig-width: 10
#| fig-height: 5

q2_arima_fc |>
  autoplot(print_retail_act, level = c(80, 95)) +
  labs(title = "Twenty-four month ARIMA forecast",
       subtitle = "Final model: ARIMA(3,1,1)(1,1,1)[12] with Box-Cox transformation",
       x = "Month",
       y = "Turnover (Millions of Dollars)")
```

@fig-q2-arima-forecast-full shows the 24-month forecasts from the final ARIMA(3,1,1)(1,1,1)$_{12}$ model in the context of the full series. The forecasts look reasonable because they continue the recent lower-turnover period, which is consistent with the long-run downward movement seen in the later part of the series.

```{r}
#| label: fig-q2-arima-forecast-zoom
#| fig-cap: "Zoomed twenty-four month forecasts from the final ARIMA model."
#| fig-width: 10
#| fig-height: 5

q2_arima_fc |>
  autoplot(print_retail_act |> filter(year(Month) >= 2018),
           level = c(80, 95)) +
  geom_vline(xintercept = forecast_start,
             linetype = "dashed") +
  labs(title = "Zoomed twenty-four month ARIMA forecast",
       subtitle = "Dashed line marks the start of the forecast period",
       x = "Month",
       y = "Turnover (Millions of Dollars)")
```

@fig-q2-arima-forecast-zoom gives a clearer view of the recent data and forecast horizon. The point forecasts continue from the latest observed level and retain strong monthly seasonality, with expected seasonal peaks and troughs.

The prediction intervals widen as the forecast horizon increases, reflecting greater uncertainty further into the future. Since the final ARIMA model still failed the Ljung–Box test, the forecasts should be interpreted cautiously. The model is the best ARIMA choice among those considered based on AICc, but the remaining residual autocorrelation suggests that the forecast intervals may not fully capture all uncertainty.

\newpage
## Question 3

*On the same graph, plot the updated data together with forecasts from all the 'best' alternative methods/models you have studied throughout the semester. Generate as many steps ahead forecasts required to cover the new data. Think about visualisation. Visually inspect these and comment on any similarities, differences, anomalies you may observe. (No more than 100 words). (12 marks)*

ETS, STL, and ARIMA were retained from earlier best models. SNAIVE and TSLM serve as baselines representing seasonal persistence and deterministic trends. All models were fitted only to data up to June 2023.

```{r}
# Training set = data available at the time of GA2/GA3 (up to 2023 Jun)
train_data <- print_retail_act |> 
  filter(Month <= yearmonth("2023 Jun"))

# Define the transformation
lambda_train <- train_data |>
  features(y, guerrero) |>
  pull(lambda_guerrero)

# Refit the models
fit_all <- train_data |> 
  model(
    `Benchmark (SNAIVE)` = SNAIVE(y),
    `Regression (TSLM)` = TSLM(box_cox(y, lambda_train) ~ trend() + season()),
    `STL Decomposition` = decomposition_model(
      STL(box_cox(y, lambda_train) ~ season(window = "periodic")),
      NAIVE(season_adjust)),
    `Best ETS (M,N,M)` = ETS(y ~ error("M") + trend("N") + season("M")),
    `Best ARIMA` = ARIMA(box_cox(y, lambda_train) ~ pdq(3,1,1) + PDQ(1,1,1)))

fc_all <- fit_all |> 
  forecast(h = "24 months")
```

```{r}
#| label: fig-q3-all-forecasts-zoom
#| fig-cap: "Updated turnover series to June 2025 with 24-month-ahead forecasts from selected models fitted on data up to June 2023. The dashed vertical line marks the start of the updated data period."

fc_all |> 
  autoplot(print_retail_act |> filter_index("2021 Jan" ~ .), 
           level = NULL, linewidth = 0.8) +
  geom_vline(xintercept = yearmonth("2023 Jul"),
             linetype = "dashed",
             colour = "black",
             linewidth = 0.8) +
  labs(title = "Forecast comparison using models fitted to June 2023",
       subtitle = "Dashed line marks the start of the updated data period",
       y = "Turnover (Millions of Dollars)", x = "Month", colour = "Model") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

**Visual Inspection and Commentary:**

@fig-q3-all-forecasts-zoom shows that the main similarity is that all models retain some seasonal movement. However, they differ in level and peak size. TSLM clearly over-forecasts the updated period, while STL gives the closest overall forecasts because it stays near the observed data for most months. ETS and SNAIVE capture the timing of the December seasonal peaks relatively well, but still under-forecast the peak values. ARIMA performs reasonably in 2023 but under-forecasts more clearly by late 2024. The sharp December 2024 peak is the main anomaly, as none of the models forecast it well.

\newpage
## Question 4

*Summarise the accuracy of the competing forecasts over the new/updated data and choose the best two models/methods from this evaluation. (6 marks)*

```{r}

```

\newpage
## Question 5

*Use the best two models and generate forecasts for your series based on the end of your updated sample. Plot these (think about visualisation). Comment on these. (No more than 100 words). (6 marks)*

```{r}

```

\newpage
## Question 6

*You have now applied a range of modelling frameworks throughout the semester and developed forecasts for multiple time series. For this final task, write a short report (300–500 words) reflecting on your thoughts and experiences with these frameworks. Your report should address aspects such as: which methods worked best for the series you analysed; which framework did you find the easiest to understand and implement; what were the key strengths and limitations of each approach. You should consider including at least one comment on each of: benchmarks,  decomposition, ETS, ARIMA and regression frameworks. Additionally, reflect on whether a regression-based model might be more suitable for forecasting retail data affected by COVID-19, and explain how you would approach this implementation. (20 marks)*

*This task is intentionally open-ended to give you the opportunity to think critically and share your personal insights. Go beyond the questions above if you can, consider what you enjoyed, what you didn't, what challenged you, and what you’ve learnt from all your hard work. Treat this as a report advising/informing a client (who knows statistics) while you were working for your forecasting consulting group. Include your consulting group name.*

*Congratulations on completing all tasks for this unit. I hope you’ve enjoyed the learning journey and gained valuable forecasting skills. Well done on all your hard work!*

*Cheers, George and team.*

```{r}

```