---
title: "ETF5231 Group Assignment 4"
subtitle: 'Group7 GenieUS'
author:
  - Ting-Ting Wu (35717343)
  - Tzu-Hsuan, Yang (35461586)
  - Nutchanok Saitin (35493011)
  - Heng-Hsieh, Chang (35567201)
format: pdf
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=TRUE, error=TRUE, cache=TRUE)
# write the appropriate code to read in your group data from Group_Data.csv

library(fpp3)
library(readr)
library(dplyr)
library(ggplot2)
library(patchwork)
library(knitr)
library(kableExtra)
library(readxl)
```

```{r}
#| include: false

meta <- read_csv("GA_data.csv", col_names = TRUE, n_max = 3)

dat <- read_csv("GA_data.csv", 
                col_names = colnames(meta),
                skip = 4)

print_retail_act <- dat |>
  rename(Month = "Student ID", y ="Group 7") |>
  select(Month, y) |>
  mutate(Month=yearmonth(Month)) |>
  as_tsibble(index = Month)
```

## Question 1

*Update your data from the ABS website and plot it. Do you see a COVID-19 effect? (3 marks)*

```{r}
#| label: update-dataprep

extended_dat <- read_excel("GA4_data.xlsx", sheet = "Data1", skip = 9) 

extended_series <- extended_dat |>
  rename(Month = "Series ID", y = "A3349773T") |>
  select(Month, y) |>
  mutate(Month = yearmonth(Month)) |>
  as_tsibble(index = Month)
```

```{r}
#| label: fig-extended-series
#| fig-width: 12
#| fig-height: 4
#| fig-cap: "Updated turnover series from Apr 1982 to Jun 2025. The shaded area highlights March 2020 to June 2022, when ABS did not publish trend estimates due to COVID-related disruption and volatility. The left plot shows the whole series, while the right plot zooms in from Jan 2018 to show recent movements."

plot_whole_series <- extended_series |>
  autoplot(y) +
  annotate(
    "rect",
    xmin = yearmonth("2020 Mar"),
    xmax = yearmonth("2022 Jun"),
    ymin = -Inf,
    ymax = Inf,
    alpha = 0.2,
    fill = "grey30"
  ) +
  labs(
    x = "Month",
    y = "Turnover (Millions of Dollars)"
  ) +
    theme_minimal()

plot_zoomed <- extended_series |>
  filter(Month >= yearmonth("2018 Jan")) |>
  autoplot(y) +
  annotate(
    "rect",
    xmin = yearmonth("2020 Mar"),
    xmax = yearmonth("2022 Jun"),
    ymin = -Inf,
    ymax = Inf,
    alpha = 0.2,
    fill = "grey30"
  ) +
  labs(
    x = "Month",
    y = "Turnover (Millions of Dollars)"
  ) +
    theme_minimal()

(plot_whole_series | plot_zoomed) +
  plot_annotation(
    title = "Turnover Series with COVID-19 Period Highlighted",
    subtitle = "Newspaper and book retailing, Australian Capital Territory"
  )
```

@fig-extended-series shows a long-term decline in turnover before COVID-19. During the COVID-disruption period, turnover remained low and volatile, with some sharp temporary movements, including the lowest point in the series. After the COVID-disruption period, turnover shows a slight recovery, but remains at a relatively low level, with the seasonal pattern still visible.

```{r}
#| label: fig-season-plot
#| fig-height: 6

extended_series |>
  filter(Month >= yearmonth("2018 Jan")) |>
  gg_season(y, labels = "both") +
  labs(
    title = "Turnover (Jan 2018 - Jun 2025)",
    subtitle = "Newspaper and book retailing, Australian Capital Territory",
    x = "Month",
    y = "Turnover (Millions of Dollars)"
  ) +
    theme_minimal()
```

@fig-season-plot shows that turnover generally declined over time while retaining a consistent seasonal pattern, with higher values toward the end of the year, especially in December. During the COVID-disruption period, the 2021 line falls sharply from July to October and reaches the lowest point in September, suggesting a strong temporary weakening. After the COVID-disruption period, turnover increases slightly in some months, but remains relatively low compared with 2018–2020.

\newpage
## Question 2

*Identify the best possible ARIMA model for your data and generate forecasts. Present a complete analysis and justifications for each step you take. There is no word limit here. No more hints will be provided - we leave it up to you to provide a full analysis. Treat this as a consulting job for a client wanting some ARIMA forecasts (20 marks)*

```{r}

```

\newpage
## Question 3

*On the same graph, plot the updated data together with forecasts from all the 'best' alternative methods/models you have studied throughout the semester. Generate as many steps ahead forecasts required to cover the new data. Think about visualisation. Visually inspect these and comment on any similarities, differences, anomalies you may observe. (No more than 100 words). (12 marks)*

```{r}

```

\newpage
## Question 4

*Summarise the accuracy of the competing forecasts over the new/updated data and choose the best two models/methods from this evaluation. (6 marks)*

```{r}

```

\newpage
## Question 5

*Use the best two models and generate forecasts for your series based on the end of your updated sample. Plot these (think about visualisation). Comment on these. (No more than 100 words). (6 marks)*

```{r}

```

\newpage
## Question 6

*You have now applied a range of modelling frameworks throughout the semester and developed forecasts for multiple time series. For this final task, write a short report (300–500 words) reflecting on your thoughts and experiences with these frameworks. Your report should address aspects such as: which methods worked best for the series you analysed; which framework did you find the easiest to understand and implement; what were the key strengths and limitations of each approach. You should consider including at least one comment on each of: benchmarks,  decomposition, ETS, ARIMA and regression frameworks. Additionally, reflect on whether a regression-based model might be more suitable for forecasting retail data affected by COVID-19, and explain how you would approach this implementation. (20 marks)*

*This task is intentionally open-ended to give you the opportunity to think critically and share your personal insights. Go beyond the questions above if you can, consider what you enjoyed, what you didn't, what challenged you, and what you’ve learnt from all your hard work. Treat this as a report advising/informing a client (who knows statistics) while you were working for your forecasting consulting group. Include your consulting group name.*

*Congratulations on completing all tasks for this unit. I hope you’ve enjoyed the learning journey and gained valuable forecasting skills. Well done on all your hard work!*

*Cheers, George and team.*

```{r}

```